\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-11-01 standardmodel.tex
Save. Changing Notation.

Note that, since every piece is intrinsic well-typed, $\denotesS{t}$ means the same as $\denotesS{\goodTerm{\Gamma}{t}{T}}$ (i.e. every term is tracking its context and type).\footnote{And our (pseudo) Agda formulation is well-typed as well. Latex-formulation should be well-typed but we discarded all the type information in the paper.}

@@

  \denotesS{\goodCtx{\Gamma}{i}} & \text{ is a Set and we define } \goodCtxS{\Gamma}{i} :\iff \denotesS{\goodCtx{\Gamma}{i}} \text{ is defined}\tikzmarkend{stmod1} 

@@

  \denotesS{\goodCtx{\Gamma, A}{i}} &= \{(\gamma, a) : \gamma \in \denotesS{\goodCtx{\Gamma}{}}, a \in \denotesS{\goodType{\Gamma}{A}{}}(\gamma)\}

@@

  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod2}(0.05,-0.15)(-0.10,0.30) \denotesS{\goodSub{\Gamma}{\_}{\Delta}} & \text{ is a Set and we define }  {\goodSubS{\Gamma}{\gamma}{\Delta}} : \iff \denotesS{\gamma} \in \denotesS{\goodSub{\Gamma}{\_}{\Delta}} \tikzmarkend{stmod2} 

@@

  \denotesS{\goodSub{\Gamma}{\_}{\Delta}} & = \denotesS{\goodCtx{\Gamma}{i}} \rightarrow \denotesS{\goodCtx{\Delta}{j}} \\
  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod3}(0.05,-0.15)(-0.10,0.30) \denotesS{\goodTerm{\Gamma}{\_}{T}} & \text{ is a Set (of dependent function) } \\& \text{and we define }  {\goodTermS{\Gamma}{t}{T}} : \iff \denotesS{t} \in \denotesS{\goodTerm{\Gamma}{\_}{T}}

@@

  \denotesS{\goodTerm{\Gamma}{\_}{T}} & = \prod_{\gamma \in \denotesS{\goodCtx{\Gamma}{i}}}\denotesS{T}(\gamma) \\

@@

  \denotesS{\goodType{\Gamma}{T}{i}} & \text{ is a family of sets indexed by } \denotesS{\goodCtx{\Gamma}{i}}  

@@

  & \text{ and we define } \goodTypeS{\Gamma}{T}{i} : \iff \denotesS{\goodType{\Gamma}{T}{i}} \text{ is defined } \tikzmarkend{stmod4} 

@@

  \denotesS{\goodType{\Gamma}{\Sigma A B}{}}(\gamma) &:= \sum_{a \in \denotesS{A}\gamma} \denotesS{B}(\gamma, a) \\

@@

  \denotesS{\goodSig{\Gamma}{\_}{n}} & \text{ is a Set  and we define  } {\goodSigS{\Gamma}{\sigma}{n}} :\iff \denotesS{\sigma} \in  \denotesS{\goodSig{\Gamma}{\_}{n}} \tikzmarkend{stmod5} 

@@

  \denotesS{\goodType{\Gamma}{\cC \sigma}{}} & \text{ is a family of sets, defined inductively on } \sigma \tikzmarkend{stmod6} \\   
  \denotesS{\goodSig{\Gamma}{\_}{n+1}} &= \{

@@

      {\goodSigS{\Gamma}{\sigma}{n}}
      \land  {\goodTypeS{\Gamma}{A}{}} \\  
      & \quad \quad \quad \land  {\goodSealS{\Gamma}{f}{\sigma}{A}}
      \land  {\goodTypeS{\Gamma, A}{T}{i}} 

@@

  \denotesS{\goodSig{\Gamma}{\_}{0}} &= \{\star\} \quad \text{ a specific singleton set} \\
  \denotesS{\goodType{\Gamma}{\cC \sigma}{}} &= \denotesS{\Sigma} \denotesS{\goodType{\Gamma}{\cC \tau}{}} \denotesS{T["sf" \ f]} \\
  & \quad \text{when } \denotesS{\sigma} = (\denotesS{\tau}, \denotesS{A}, \denotesS{f} , \denotesS{T}) \in \denotesS{\goodSig{\Gamma}{\_}{n+1}} \\
  \denotesS{\goodType{\Gamma}{\cC \sigma}{i}}&= \denotesS{\top}\quad \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gamma}{\_}{0}} \\

@@

  \denotesS{\goodType{\Gamma}{\cL \sigma}{i}} & \text{ is a family of sets, defined inductively on } \sigma \tikzmarkend{stmod7}  \\
  \denotesS{\goodType{\Gamma}{\cL \sigma}{i}} (\gamma)&= \{(m, t) :  m \in \denotesS{\cL \tau}(\gamma) \land t \in \prod_{m' \in \denotesS{A}(\gamma)}\denotesS{T}(\gamma, m')  \}  \\
  & \quad \text{when } \denotesS{\sigma} = (\denotesS{\tau}, \denotesS{A}, \_, \denotesS{T}) \in \denotesS{\goodSig{\Gamma}{\_}{n+1}} \\
  \denotesS{\goodType{\Gamma}{\cL \sigma}{i}} (\gamma)&= \{\star\} \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gamma}{\_}{0}} \\
  \denotesS{\goodSig{\Gamma}{\nu\cdot}{0}} &= \star \\
  \denotesS{\goodSig{\Gamma}{\nu^+ \ \sigma \ f \ T}{n+1}} &= (\sigma, A, f, T) \\
  \denotesS{\goodTerm{\Gamma}{\mu\cdot}{\cL \nu\cdot} } (\gamma) &= \star \\
  \denotesS{\goodTerm{\Gamma}{\mu^+ \ o \ t}{\cL(\nu^+ \ \sigma  \ T)}} \ (\gamma)&= (o(\gamma), (\lambda \ m' . t'(\gamma, m'))) \\
  \denotesS{\goodTerm{\Gamma}{p_1\mu~t}{\cL(p_1\nu~\sigma)}}(\gamma) &=  "fst"~(t\gamma) \\ 

@@

  \denotesS{\goodTerm{\Gamma}{\cCt \{\sigma\} o}{\cC\sigma}} & \text{ is a dependent function, defined inductively on } {\sigma} 

@@

  \denotesS{\goodTerm{\Gamma}{\cCt o}{\cC\sigma}}
  &= \denotesS{{\cCt (p_1\mu \ o)}}, \denotesS{(p_2\mu \ o)["sf" \ f][(id, \cCt (p_1\mu \ o))]} \\
  & \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gamma}{\_}{n+1}} \\
  \denotesS{\goodTerm{\Gamma}{\cCt o}{\cC\sigma}}
  &= \star \quad \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gamma}{\_}{0}} \\
  \denotesS{\goodType{\Gamma}{\bot}{i}} \ (\gamma) &= \emptyset \\
  \denotesS{\goodType{\Gamma}{\cB}{i}} \ (\gamma) &= \{0, 1\} \quad \text{a specific two element set} \\

@@

  \denotesS{\goodWSig{\Gamma}{\_}{n}} &\text{ is a set of list of pairs of type interpretation of length } n \\
  & \text{ and we define } \goodWSigS{\Gamma}{\tau}{n} :\iff \denotesS{\goodWSig{\Gamma}{\tau}{n}} \in \denotesS{\goodWSig{\Gamma}{\_}{n}}

@@

  \denotesS{\goodWSig{\Gamma}{w\cdot}{0}} &= [] \quad \text{ an empty list} \\ 
  \denotesS{\goodWSig{\Gamma}{(w^+ \ A \ B \ \tau)}{n+1}}
  &= (\denotesS{A}, \denotesS{B})~"::"~\denotesS{\goodWSig{\Gamma}{\tau}{n}} \\ 
  \denotesS{\goodType{\Gamma}{\pi^j_1~\tau}{}} &= "fst"~({\denotesS{\tau}}_{j})\\
  \denotesS{\goodType{\Gamma, \pi^j_1~\tau}{\pi^j_2~\tau}{}} &= "snd"~({\denotesS{\tau}}_j)\\

@@

We encourage the reader to read the (pseudo-)Agda-style proof with more type annotation so it is clearer about the mathematical object doing the interpretation. Here, we need to clarify some ambiguity in our latex formulation. For example, when we wrote ``$\denotesS{\goodType{\Gamma}{\cC \sigma}{}}$ defined inductively on ${\sigma}$'', the best way to understand it is to consider we are defining a function $\denotesS{\goodType{\Gamma}{\cC \_}{}}$ as the interpretation for $\goodType{\Gamma}{\cC \_}{}$ that mapping $\denotesS{\sigma}$ to $\denotesS{\goodType{\Gamma}{\cC \sigma}{}}$, inductively on the length of the signature. The case is similar for $\denotesS{\cL \_}$, $\denotesS{\cCt \_}$ and  $\denotesS{\Sigma \ \_ \ \_}$. \textbf{This is actually the formulation in Agda}---QIIT syntax consider $\cC : "Sig"~\Gamma~n \to "Ty"~\Gamma$ as a constructor and our model needs to interpret this constructor to a function $\cC^S : "Sig"^S \Gamma^S~n \to "Ty"^S~\Gamma^S$ with each judgement also interpreted.

@@

Note that $\denotesS{\goodSig{\Gamma}{s}{n}}$ and $\denotesS{\goodType{\Gamma}{\cC\sigma}{i}}$ is mutual recursively defined together---because \\ $\goodSealS{\Gamma}{f}{\sigma}{A}$ relies on $\cC$.

@@

The verification of validity of this model is largely omitted here and we refer the reader to the appendix. To verify our model preserve the judgemental equality, usually unpacking the definition is enough---for example, to verify $\beta$-rule for linkage projection, we only need to see that $\denotesS{\goodTerm{\Gamma}{p_1\mu~(\mu^+~o~t)}{\cL\sigma}} = \denotesS{\goodTerm{\Gamma}{o}{\cL\sigma}}$ by function extensionality. One exception is for $\cC$ and $\goodSig{\_}{\_}{n}$, the substitution law around these two are also proved together by induction because they were mutual-recursively defined together. The compatibility lemmas can be verified directly---for example $\frac{\goodSigS{\Gamma}{\sigma}{n+1}\quad\goodTermS{\Gamma}{o}{\cL\sigma}}{\goodTermS{\Gamma}{(p_1\mu~o)}{\cL(p_1\nu~\sigma)}}$ is proved by simply unfolding the definition.

>>>>

Note that, since every piece is intrinsic well-typed, $\denotesS{t}$ means the same as $\denotesS{\goodTerm{\Gm}{t}{T}}$ (i.e. every term is tracking its context and type).\footnote{And our (pseudo) Agda formulation is well-typed as well. Latex-formulation should be well-typed but we discarded all the type information in the paper.}

@@

  \denotesS{\goodCtx{\Gm}{i}} & \text{ is a Set and we define } \goodCtxS{\Gm}{i} :\iff \denotesS{\goodCtx{\Gm}{i}} \text{ is defined}\tikzmarkend{stmod1} 

@@

  \denotesS{\goodCtx{\Gm, A}{i}} &= \{(\Gm, a) : \Gm \in \denotesS{\goodCtx{\Gm}{}}, a \in \denotesS{\goodType{\Gm}{A}{}}(\Gm)\}

@@

  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod2}(0.05,-0.15)(-0.10,0.30) \denotesS{\goodSub{\Gm}{\_}{\Delta}} & \text{ is a Set and we define }  {\goodSubS{\Gm}{\Gm}{\Delta}} : \iff \denotesS{\Gm} \in \denotesS{\goodSub{\Gm}{\_}{\Delta}} \tikzmarkend{stmod2} 

@@

  \denotesS{\goodSub{\Gm}{\_}{\Delta}} & = \denotesS{\goodCtx{\Gm}{i}} \rightarrow \denotesS{\goodCtx{\Delta}{j}} \\
  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod3}(0.05,-0.15)(-0.10,0.30) \denotesS{\goodTerm{\Gm}{\_}{T}} & \text{ is a Set (of dependent function) } \\& \text{and we define }  {\goodTermS{\Gm}{t}{T}} : \iff \denotesS{t} \in \denotesS{\goodTerm{\Gm}{\_}{T}}

@@

  \denotesS{\goodTerm{\Gm}{\_}{T}} & = \prod_{\Gm \in \denotesS{\goodCtx{\Gm}{i}}}\denotesS{T}(\Gm) \\

@@

  \denotesS{\goodType{\Gm}{T}{i}} & \text{ is a family of sets indexed by } \denotesS{\goodCtx{\Gm}{i}}  

@@

  & \text{ and we define } \goodTypeS{\Gm}{T}{i} : \iff \denotesS{\goodType{\Gm}{T}{i}} \text{ is defined } \tikzmarkend{stmod4} 

@@

  \denotesS{\goodType{\Gm}{\Sigma A B}{}}(\Gm) &:= \sum_{a \in \denotesS{A}\Gm} \denotesS{B}(\Gm, a) \\

@@

  \denotesS{\goodSig{\Gm}{\_}{n}} & \text{ is a Set  and we define  } {\goodSigS{\Gm}{\sigma}{n}} :\iff \denotesS{\sigma} \in  \denotesS{\goodSig{\Gm}{\_}{n}} \tikzmarkend{stmod5} 

@@

  \denotesS{\goodType{\Gm}{\TyTkg {\sigma}}{}} & \text{ is a family of sets, defined inductively on } \sigma \tikzmarkend{stmod6} \\   
  \denotesS{\goodSig{\Gm}{\_}{n+1}} &= \{

@@

      {\goodSigS{\Gm}{\sigma}{n}}
      \land  {\goodTypeS{\Gm}{A}{}} \\  
      & \quad \quad \quad \land  {\goodSealS{\Gm}{f}{\sigma}{A}}
      \land  {\goodTypeS{\Gm, A}{T}{i}} 

@@

  \denotesS{\goodSig{\Gm}{\_}{0}} &= \{\star\} \quad \text{ a specific singleton set} \\
  \denotesS{\goodType{\Gm}{\TyTkg {\sigma}}{}} &= \denotesS{\Sigma} \denotesS{\goodType{\Gm}{\TyTkg {\tau}}{}} \denotesS{T["sf" \ f]} \\
  & \quad \text{when } \denotesS{\sigma} = (\denotesS{\tau}, \denotesS{A}, \denotesS{f} , \denotesS{T}) \in \denotesS{\goodSig{\Gm}{\_}{n+1}} \\
  \denotesS{\goodType{\Gm}{\TyTkg {\sigma}}{i}}&= \denotesS{\top}\quad \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{0}} \\

@@

  \denotesS{\goodType{\Gm}{\cL \sigma}{i}} & \text{ is a family of sets, defined inductively on } \sigma \tikzmarkend{stmod7}  \\
  \denotesS{\goodType{\Gm}{\cL \sigma}{i}} (\Gm)&= \{(m, t) :  m \in \denotesS{\cL \tau}(\Gm) \land t \in \prod_{m' \in \denotesS{A}(\Gm)}\denotesS{T}(\Gm, m')  \}  \\
  & \quad \text{when } \denotesS{\sigma} = (\denotesS{\tau}, \denotesS{A}, \_, \denotesS{T}) \in \denotesS{\goodSig{\Gm}{\_}{n+1}} \\
  \denotesS{\goodType{\Gm}{\cL \sigma}{i}} (\Gm)&= \{\star\} \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{0}} \\
  \denotesS{\goodSig{\Gm}{\nu\cdot}{0}} &= \star \\
  \denotesS{\goodSig{\Gm}{\nu^+ \ \sigma \ f \ T}{n+1}} &= (\sigma, A, f, T) \\
  \denotesS{\goodTerm{\Gm}{\mu\cdot}{\cL \nu\cdot} } (\Gm) &= \star \\
  \denotesS{\goodTerm{\Gm}{\mu^+ \ o \ t}{\cL(\nu^+ \ \sigma  \ T)}} \ (\Gm)&= (o(\Gm), (\lambda \ m' . t'(\Gm, m'))) \\
  \denotesS{\goodTerm{\Gm}{p_1\mu~t}{\cL(p_1\nu~\sigma)}}(\Gm) &=  "fst"~(t\Gm) \\ 

@@

  \denotesS{\goodTerm{\Gm}{\Tkg {o}}{\TyTkg{\sigma}}} & \text{ is a dependent function, defined inductively on } {\sigma} 

@@

  \denotesS{\goodTerm{\Gm}{\Tkg {o}}{\TyTkg{\sigma}}}
  &= \denotesS{{\Tkg {p_1\mu \ o}}}, \denotesS{(p_2\mu \ o)["sf" \ f][(id, \Tkg {p_1\mu \ o})]} \\
  & \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{n+1}} \\
  \denotesS{\goodTerm{\Gm}{\Tkg {o}}{\TyTkg{\sigma}}}
  &= \star \quad \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{0}} \\
  \denotesS{\goodType{\Gm}{\bot}{i}} \ (\Gm) &= \emptyset \\
  \denotesS{\goodType{\Gm}{\cB}{i}} \ (\Gm) &= \{0, 1\} \quad \text{a specific two element set} \\

@@

  \denotesS{\goodWSig{\Gm}{\_}{n}} &\text{ is a set of list of pairs of type interpretation of length } n \\
  & \text{ and we define } \goodWSigS{\Gm}{\tau}{n} :\iff \denotesS{\goodWSig{\Gm}{\tau}{n}} \in \denotesS{\goodWSig{\Gm}{\_}{n}}

@@

  \denotesS{\goodWSig{\Gm}{w\cdot}{0}} &= [] \quad \text{ an empty list} \\ 
  \denotesS{\goodWSig{\Gm}{(w^+ \ A \ B \ \tau)}{n+1}}
  &= (\denotesS{A}, \denotesS{B})~"::"~\denotesS{\goodWSig{\Gm}{\tau}{n}} \\ 
  \denotesS{\goodType{\Gm}{\pi^j_1~\tau}{}} &= "fst"~({\denotesS{\tau}}_{j})\\
  \denotesS{\goodType{\Gm, \pi^j_1~\tau}{\pi^j_2~\tau}{}} &= "snd"~({\denotesS{\tau}}_j)\\

@@

We encourage the reader to read the (pseudo-)Agda-style proof with more type annotation so it is clearer about the mathematical object doing the interpretation. Here, we need to clarify some ambiguity in our latex formulation. For example, when we wrote ``$\denotesS{\goodType{\Gm}{\TyTkg {\sigma}}{}}$ defined inductively on ${\sigma}$'', the best way to understand it is to consider we are defining a function $\denotesS{\goodType{\Gm}{\TyTkg {\_}}{}}$ as the interpretation for $\goodType{\Gm}{\TyTkg {\_}}{}$ that mapping $\denotesS{\sigma}$ to $\denotesS{\goodType{\Gm}{\TyTkg {\sigma}}{}}$, inductively on the length of the signature. The case is similar for $\denotesS{\cL \_}$, $\denotesS{\Tkg {\_}}$ and  $\denotesS{\Sigma \ \_ \ \_}$. \textbf{This is actually the formulation in Agda}---QIIT syntax consider $\TyTkg{} : "Sig"~\Gm~n \to "Ty"~\Gm$ as a constructor and our model needs to interpret this constructor to a function $\TyTkg{}^S : "Sig"^S \Gm^S~n \to "Ty"^S~\Gm^S$ with each judgement also interpreted.

@@

Note that $\denotesS{\goodSig{\Gm}{s}{n}}$ and $\denotesS{\goodType{\Gm}{\TyTkg{\sigma}}{i}}$ is mutual recursively defined together---because \\ $\goodSealS{\Gm}{f}{\sigma}{A}$ relies on $\TyTkg{}$.

@@

The verification of validity of this model is largely omitted here and we refer the reader to the appendix. To verify our model preserve the judgemental equality, usually unpacking the definition is enough---for example, to verify $\beta$-rule for linkage projection, we only need to see that $\denotesS{\goodTerm{\Gm}{p_1\mu~(\mu^+~o~t)}{\cL\sigma}} = \denotesS{\goodTerm{\Gm}{o}{\cL\sigma}}$ by function extensionality. One exception is for $\TyTkg{}$ and $\goodSig{\_}{\_}{n}$, the substitution law around these two are also proved together by induction because they were mutual-recursively defined together. The compatibility lemmas can be verified directly---for example $\frac{\goodSigS{\Gm}{\sigma}{n+1}\quad\goodTermS{\Gm}{o}{\cL\sigma}}{\goodTermS{\Gm}{(p_1\mu~o)}{\cL(p_1\nu~\sigma)}}$ is proved by simply unfolding the definition.
